The Causes of Damage in Toronto Fire Incidents

### 2024-09-27 paper/paper.qmd
Step4: Clean unused files, add title

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - David Qi
thanks: "Code and data are available at: https://github.com/UTDQi/starter_folder"
date: today
date-format: long
abstract: "This paper discusses some basic findings around the monetary and lives loss in Toronto fire incidents. By using data from open data Toronto and choosing the relevant entries, we have identified possible factors that influences monetary and lives loss in fire incidents. "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

@@

ggplot(top_fire_causes, aes(x = reorder(Possible_Cause, -Count), y = Avg_Civilian_Casualties)) +
  geom_bar(stat = "identity", fill = "orange", width = 0.6) +  # Adjust bar width
  geom_text(aes(label = round(Avg_Civilian_Casualties, 3)), size = 4, color = "black") +  # Add average casualties labels
  theme_minimal(base_size = 15) +  # Increase base font size for readability
  labs(title = "Average Civilian Casualties for Top 5 Causes",
       x = "Possible Cause",
       y = "Average Casualties") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),  # Rotate x-axis labels
        axis.text.y = element_text(size = 6),  # Increase y-axis text size
        plot.margin = margin(t = 20, r = 20, b = 20, l = 20),  # Add margins
        panel.grid.major.x = element_blank())  # Remove grid lines
```

@@

>>>>

---
title: "The Causes of Damage in Toronto Fire Incidents"
author: 
  - David Qi
thanks: "Code and data are available at: https://github.com/UTDQi/Fire_Incidents"
date: today
date-format: long
abstract: "This paper discusses some basic findings around the monetary and lives loss in Toronto fire incidents. By using data from open data Toronto and choosing the relevant entries, we have identified possible factors that influences monetary and lives loss in fire incidents. We identified unattended ignition sources as a significant risk factor for civilian casualties, highlighting the need for public awareness campaigns to educate communities about these potential hazards and noted that fires with undetermined causes resulted in greater property loss. Interestingly, we observed that increasing the number of responding personnel beyond 90 led to a reduction in civilian casualties, which contrasts with our initial expectations."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

It's important to note that a relatively low average damage does not imply that a fire cause is insignificant. For example, improperly discarded ignition sources may be relatively less damaging according to the figures. But it is the second most frequent single cause of fires Incidents. This category has contributed to over $20 million in losses and resulted in more than 110 civilian casualties. Such statistics indicates the need for proactive measures against seemingly less damaging causes, such as public education campaign or fines for improperly discarded cigarette. As these fire also are severe threats for public safety.

@@

ggplot(top_fire_causes, aes(x = reorder(Possible_Cause, -Count), y = Avg_Civilian_Casualties)) +
  geom_bar(stat = "identity", fill = "orange", width = 0.6) +  # Adjust bar width
  geom_text(aes(label = round(Avg_Civilian_Casualties, 3)), size = 4, color = "black") +  # Add average casualties labels
  theme_minimal(base_size = 15) +  # Increase base font size for readability
  labs(title = "Average Civilian Casualties for Top 5 Causes",
       x = "Possible Cause",
       y = "Average Casualties") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),  # Rotate x-axis labels
        axis.text.y = element_text(size = 6),  # Increase y-axis text size
        plot.margin = margin(t = 20, r = 20, b = 20, l = 20),  # Add margins
        panel.grid.major.x = element_blank())  # Remove grid lines
```
\newpage
## Number of responding personnel

@@

  Our study have shown that increasing the number of responding personnel doesn't substantially decrease property loss as the effect it has on civilian causalities. While we observed that a higher number of personnel reduces civilian casualties, the same effect is not observed in property loss. This distinction suggests that the effectiveness of firefighting operations may be more closely related to the effectiveness of each person, the ability of each person at sight, rather than number of person present. Future studies should explore the underlying reasons for this difference in relationship.
